A thesis draft

### 2018-12-03 res/sections/15-ArchitecturalImplementation.tex
Fix typos

First we started assesing the Docker Swarm orchestrator. Docker Swarm is a
product directly made by the Docker Inc., so it seamlessly integrates with other
Docker tools such as Docker Compose. The installation and set-up is pretty
straightforward, and the configuration is almost automatically handled by Docker
itself. The minimum number of nodes needed to create a proper Docker Swarm
cluster is three, where one must be a manager node (designed to orchestrate the
other nodes in the cluster), while the others can be simple workers. Multiple
managers can coexist in the same cluster, as long as only one effectively
operates. The others act like backup managers in case the one orchestrating goes
down for some reason. The networking is given out-of-the box, and the same holds
for distributed persistence. While this for most of the uses cases can be
considered a pro, in our specific situation the impossibility to tweak the
network configuration was a cons, since it was important for us to have the full
control over the network, especially regarding the SFC implementation, explained
better in Chapter~\ref{chap:vnf_ns_impl}.

@@

Kubernetes does not require a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
For the initial deployment we relied on an internal Openstack instance installed
at the University of Padova made of four nodes with 4GB of RAM, 10GB of SSD
storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
7). We performed the installation using the official Kubernetes repositories,
taking care of performing a accurated cluster configuration, as better explained
in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

Terminated the Harbor development, we ended up in a situation similar the one
depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
It is possible to see how the Openbaton framework is not necessary anymore:
Harbor, in fact, ended up substituting it. We then started to think about the
challenges of integrating the SFC structure inside a Kubernetes cluster. We
identified the following challenges:
\begin{itemize}
\item whole system scalability: we want for every component to have the
  possibility of scaling during intense workloads, without the necessity of
  human intervention;
\item ingress and egress points: specularized entities of our framework handling
  SFC encapsulation/decapsulation session data;
\item dynamic packet routing: as formulated by one of our requirements, packet
  routing had to be dynamic. We immediately identified a possibility to create
  this functionality with the already-built features of Kubernetes;
\item a framework to automatically forward the data elaborated by the VNF: we
  identified the possibility to build a framework that automatically handled the
  packet receiving from another VNF, it properly called the function that had to
  elaborate the packet payload and then it forwarded the result to the next VNF;
\item connection status: having ingress and egress points implicates the
  necessity to save some persistent data in a distributed manner.
\end{itemize}
In the following, we provide some details on how we achieved the aforementioned
properties.

@@

\subsubsection{Keeping the connection status}
Desining for scalability gave rise to new issues. The most important is keeping
up stateful connections, since the ingress and egress points could be more than
one container. With this consideration in mind, we started thinking at how and
where keeping information about the ingress container handling the connection
with the sender $S$ and about the egress container talking with the receiver
$R$. Since the Roulette component has a complete end-to-end topology view of the
system, we decided to store these additional information (we called it
\emph{endpoints}) in it.

>>>>

First we started assessing the Docker Swarm orchestrator. Docker Swarm is a
product directly made by the Docker Inc., so it seamlessly integrates with other
Docker tools such as Docker Compose. The installation and set-up is pretty
straightforward, and the configuration is almost automatically handled by Docker
itself. The minimum number of nodes needed to create a proper Docker Swarm
cluster is three, where one must be a manager node (designed to orchestrate the
other nodes in the cluster), while the others can be simple workers. Multiple
managers can coexist in the same cluster, as long as only one effectively
operates. The others act like backup managers in case the one orchestrating goes
down for some reason. The networking is given out-of-the box, and the same holds
for distributed persistence. While this for most of the uses cases can be
considered a pro, in our specific situation the impossibility to tweak the
network configuration was a cons, since it was important for us to have the full
control over the network, especially regarding the SFC implementation, explained
better in Chapter~\ref{chap:vnf_ns_impl}.

@@

Kubernetes does not require a minimum number of nodes in order to properly work.
Nonetheless, we decided to use a four node cluster configured in the following
way: three slaves nodes (also known as \emph{minions}) and one master node. The
Kubernetes version installed was the 1.11, enabling us to exploit new features
like a reworked DNS lookup system and new storage features\footnote{A full
  kubernetes log release can be found at:
  \url{https://kubernetes.io/blog/2018/06/27/kubernetes-1.11-release-announcement/}}.
For the initial deployment we relied on an internal Openstack instance installed
at the University of Padua made of four nodes with 4GB of RAM, 10GB of SSD
storage and 2vCPU on Ubuntu 16.04 machines (that we later swapped with CentOS
7). We performed the installation using the official Kubernetes repositories,
taking care of performing an accurate cluster configuration, as better explained
in Appendix~\ref{kubernetes}. After the initial installation, we obtained a
configuration similar to the one illustrated in
Figure~\ref{chap:archimpl:sec:secondattempt:img:k8scp}. In this figure is
possible to denote different components on the master node and on the minions,
which will be described.

@@

Terminated the Harbor development, we ended up in a situation similar the one
depicted in Figure~\ref{chap:archimple:sec:secondattempt:img:attempt2v1harbor}.
It is possible to see how the Openbaton framework is not necessary anymore:
Harbor, in fact, ended up substituting it. We then started to think about the
challenges of integrating the SFC structure inside a Kubernetes cluster. We
identified the following challenges:
\begin{itemize}
\item whole system scalability: we want for every component to have the
  possibility of scaling during intense workloads, without the necessity of
  human intervention;
\item ingress and egress points: specialized entities of our framework handling
  SFC encapsulation/decapsulation session data;
\item dynamic packet routing: as formulated by one of our requirements, packet
  routing had to be dynamic. We immediately identified a possibility to create
  this functionality with the already-built features of Kubernetes;
\item a framework to automatically forward the data elaborated by the VNF: we
  identified the possibility to build a framework that automatically handled the
  packet receiving from another VNF, it properly called the function that had to
  elaborate the packet payload and then it forwarded the result to the next VNF;
\item connection status: having ingress and egress points implicates the
  necessity to save some persistent data in a distributed manner.
\end{itemize}
In the following, we provide some details on how we achieved the aforementioned
properties.

@@

\subsubsection{Keeping the connection status}
Designing for scalability gave rise to new issues. The most important is keeping
up stateful connections, since the ingress and egress points could be more than
one container. With this consideration in mind, we started thinking at how and
where keeping information about the ingress container handling the connection
with the sender $S$ and about the egress container talking with the receiver
$R$. Since the Roulette component has a complete end-to-end topology view of the
system, we decided to store these additional information (we called it
\emph{endpoints}) in it.
